Tidal Heating via Internal Gravity Waves in White Dwarf Binaries

### 2019-01-15 writeups/paper.tex
stash paper

The Euler equations for an incompressible, barotropic fluid in a uniform
gravitational field are
\begin{subequations}\label{se:fc_orig}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \md{\rho}{t} &= 0 ,\label{eq:fc_density}\\
        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + g\hat{z} &= 0.
    \end{align}
\end{subequations}
$\md{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$ is the material
derivative. $\vec{u}, \rho, P$ are the fluid dynamical variables and $-g\hat{z}$
is constant gravitational acceleration. Note that at hydrostatic equilibrium
$\pd{}{t} = 0$ we have $\vec{u} = 0, \vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and
so $P_0 = \rho_0 gH$.

We nondimensionalize by taking $H = N = \rho_{ref} = 1$ where
\begin{equation}
    N^2 \equiv g^2\p*{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
        = \frac{g}{H},
\end{equation}
is the Brunt-V\"ais\"al\"a frequency. The full numerical details of the
simulation are described in \autoref{ss:strat_impl}.

@@

As not to interfere with the incompressibility constraint, we force the system
on the density equation, replacing \autoref{eq:fc_density} with
\begin{equation}
    \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
            \cos \p*{k_{0x}x - \omega_0 t}.
\end{equation}
Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
$k_{0x}, \omega_0$ will propagate.
% Define $k_{0z}: \omega(k_{0x}, k_{0z}) = \omega_0$ the wavenumber picked
% out by $k_{0x}, \omega_0$. If $k_{0z}\sigma \ll 1$, the $\cos\p*{k_{0x}x -
% \omega_0 t}$ term permits only the $k_{0z}$ mode to propagate among the broad
% $k_z$ wavenumber spectrum permitted by the narrow Gaussian profile.

\subsection{Internal Gravity Waves}

In the small perturbation limit, we linearize \autoref{se:fc_orig} with
inhomegeneous driving term and obtain:
\begin{subequations}\label{se:vol_drive}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\vec{u}_1}{t} + \frac{\vec{\nabla}P}{\rho_0}
            + \frac{\rho_1 \vec{g}}{\rho_0}
            &= 0,\\
        \pd{\rho_1}{t} - \frac{u_{1z} \rho_0}{H}
            &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
                e^{i(k_xx - \omega t)}.
    \end{align}
\end{subequations}
We have substituted $\cos\p*{k_xx - \omega_0t}$ for a complex exponential for
easier manipulation.

It is well known that the solutions to the $F = 0$ homogeneous
\autoref{se:vol_drive} are of form (cite)
\begin{subequations}\label{se:lin_sol}
    \begin{align}
        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}}.
    \end{align}
\end{subequations}

To solve the inhomogeneous system, we may first approximate the driving term
using $e^{\frac{-(z - z_0)^2}{2\sigma^2}} \approx \sqrt{2\pi \sigma^2}\delta(z -
z_0)$ if $\sigma \to 0$. This system is solved exactly by matching the two
homogeneous solutions above and below $z_0$, We may then approximately relax the
solution to $k_{0z}\sigma \lesssim 1$. This may be solved to obtain the solution
\begin{align}
    A(F) &= \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}},\\
    u_{1z} &= A(F) \times\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}
An extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises compared to the
$\delta$-function solution when evaluating the Fourier Transform of
$e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$.

% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
% exact solution of nonlinear.
% Linear solution is globally unstable to RTI though, thus not stable
% equilibrium of IVP
%
% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
% wave-induced mean flow, horizontal momentum flux

Furthermore, in the low-amplitude, dissipation free limit, the time-averaged
total $x$-momentum flux in the $\hat{z}$ direction can be computed. Since all
dynamical variables $\propto e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and
time averaging are equivalent, so we may write
\begin{equation}
    F_{px} \equiv \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
\end{equation}
We will denote this $F_{px} = \ev*{\rho u_xu_z}_x$ for simplicity. For a given
forcing amplitude then:
\begin{equation}
    F_{px}(F) \approx -\frac{A(F)^2}{2}\rho_0(z_0) \frac{k_z}{k_x}.
        \label{eq:fpx_lin}
\end{equation}
We've approximated $k_zH \ll 1$ and noticed $\ev*{\rho_1 u_xu_z}_x = 0$ in the
above for clarity.

Finally, it may be noted that horizontally averaged horizontal momentum
$p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} = 0$, and since
flux is transported on the group velocity timescale, this simplifies to the
known expression for the wave-induced horizontal mean flow (cite)
\begin{equation}
    \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
\end{equation}
We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
velocity. In terms of our forcing amplitude,
\begin{equation}
    \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
        \label{eq:u0_lin}
\end{equation}
where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
\frac{1}{4H^2}}^{3/2}}$.

@@

@@

@@

\section{Equation Implementations}

We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
the number of spectral modes in the respective dimensions.

\subsection{Stratified Fluid}\label{ss:strat_impl}

We introduce variable $T = P/\rho$. We then mandate $\rho_0, T_0$ background
fields satisfy hydrostatic equilibrium $\vec{\nabla}T_0 + T_0 \vec{\nabla}\rho_0
+ \vec{g} = 0$. Taking isothermal stratification, we find $T_0 = gH$. Making
then substitution of variables $\Upsilon = \ln \rho - \ln \rho_0$ and $T_1 = T -
T_0$ deviations from the background state, the exact fluid equations in the new
variables are:
\begin{subequations}\label{se:fc_var}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\Upsilon}{t} - \frac{u_z}{H} &= 0,\\
        \pd{u_x}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_x
            + \pd{T}{x} + gH\pd{\Upsilon}{x}
            + T_1 \pd{\Upsilon}{x} &= 0,\\
        \pd{u_z}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_z
            + \pd{T}{z} + gH\pd{\Upsilon}{z}
            + T_1 \pd{\Upsilon}{z} - \frac{T_1}{H} &= 0.
    \end{align}
\end{subequations}

These are implemented as:
\begin{subequations}\label{se:curr_num}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\Upsilon}{t} - \frac{u_z}{H}
            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                    \cos \p*{k_xx - \omega t},\\
        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
            - \nu \nabla^2 u_x
            &= -\Gamma(z) u_x
                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
                - T_1 \pd{\Upsilon}{x},\\
        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
                - T_1 \pd{\Upsilon}{z},\\
    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
    \end{align}
\end{subequations}
where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
= 1024$, where $k_z$ is the wavenumber of the excited linear mode.

>>>>

The Euler equations for an incompressible, barotropic fluid in a uniform
gravitational field are
\begin{subequations}\label{se:fc_orig}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \md{\rho}{t} &= 0 ,\label{eq:fc_density}\\
        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho} + g\hat{z} &= 0.
    \end{align}
\end{subequations}
$\md{}{t} = \pd{}{t} + \p*{\vec{u} \cdot \vec{\nabla}}$ is the material
derivative. $\vec{u}, \rho, P$ are the fluid dynamical variables and $-g\hat{z}$
is constant gravitational acceleration. Note that at hydrostatic equilibrium
$\pd{}{t} = 0$ we have $\vec{\nabla}P_0 = -\rho_0 g\hat{z}$ and so $P_0 = \rho_0
gH$. A shear flow $u_x(x, z, t) = u_x(z)$ is permitted in hydrostatic
equilibrium, but we will initially consider no shear flow in equilibrium.

\subsection{Internal Gravity Waves: Linear Theory}

In the small perturbation limit, where flow velocities are small compared to the
characteristic space and time scales $\pd{}{t} \gg \vec{u} \cdot \vec{\nabla}$,
we linearize \autoref{se:fc_orig} by ignoring the advective components of the
material derivative, defining $\rho_1 \equiv \rho - \rho_0 \ll \rho_0$ and $P_1
\equiv P - P_0$ deviations from equilibrium and obtain:
\begin{subequations}\label{se:lin_homo}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\rho_1}{t} - \frac{u_{z} \rho_0}{H} &= 0,\\
        \pd{\vec{u}}{t} + \frac{\vec{\nabla}P_1}{\rho_0}
            + \frac{\rho_1 g\hat{z}}{\rho_0}
            &= 0.
    \end{align}
\end{subequations}

It is well known that the solutions to the $F = 0$ homogeneous
\autoref{se:lin_homo} are of form (cite)
\begin{subequations}\label{se:lin_sol}
    \begin{align}
        u_z\p*{x, z, t} &= Ae^{z/2H}e^{i(k_{0x}x + k_{0z}z - \omega_0 t)},\\
        \omega_0^2 &= \frac{N^2k_{0x}^2}{k_{0x}^2 + k_{0z}^2 + \frac{1}{4H^2}},
    \end{align}
\end{subequations}
up to undetermined amplitude $A$, where
\begin{equation}
    N^2 \equiv g^2\p*{\rd{\rho}{P} - \at{\pd{\rho}{P}}_{ad}}
        = \frac{g}{H},
\end{equation}
is the Brunt-V\"ais\"al\"a, or buoyancy, frequency.

In the weak stratification limit $k_zH \gg 1$, the solution exhibits the
following characteristics:
\begin{itemize}
    \item The amplitude of the perturbations grows like $e^{z/2H}$. Thus, the
        linear approximation is violated for sufficiently large $z$.

    \item The phase and group velocities can be computed respectively:
        \begin{align}
            \vec{c}_{ph} &= \frac{Nk_x}{k_x^2 + k_z^2 + 1/4H^2} \hat{k},\\
            \vec{c}_{g} &= \frac{N\p*{k_xk_z\hat{x}
                - \p*{k_z^2 + \frac{1}{4H^2}}\hat{z}}}
                {\p*{k_x^2 + k_z^2 + \frac{1}{4H^2}}^{3/2}} \nonumber\\
                &\approx \frac{Nk_z}{k_x^2 + k_z^2}\p*{k_x\hat{x} - k_z\hat{z}}.
        \end{align}
        We recover the usual result $\vec{c}_{ph} \cdot \vec{c}_g = 0$ (CITE).
        Note that an IGW transporting energy and momentum upwards has $k_z < 0$.

%     \item The linear solution is an exact solution to the full nonlinear
%         equations \autoref{se:fc_orig}, thanks to the advective terms vanishing
%         identically when $\vec{\nabla} \cdot \vec{u} = 0$ (CITE). However, this
%         solution is globally unstable to resonant triad interactions (CITE).

    \item The time-averaged total $x$-momentum flux in the $\hat{z}$ direction
        can be computed. Since the fluid variables have separable as $f(x, z, t)
        = f(z)e^{i(k_{0x}x - \omega_0 t)}$, $x$ averaging and time averaging are
        equivalent, so we may write ($\ev*{\dots}_x$ denotes $x$-averaging)
        \begin{align}
            F_{px, z} &\equiv \ev*{\rho u_x u_z}_x =
                \frac{1}{L_x}\int_0^{L_x}\limits \rho u_x u_z\;\mathrm{d}x
                    \nonumber\\
                &\approx -\frac{A^2}{2}\rho_{ref}\frac{k_z}{k_x}.
                    \label{eq:fpx_lin}
        \end{align}

    \item Finally, it may be noted that horizontally averaged horizontal
        momentum $p_x = \ev*{\rho u_x}_x$ obeys $\pd{p_x}{t} + \pd{F_{px}}{z} =
        0$. Since flux is transported on the group velocity timescale (CITE),
        this simplifies to the known expression for the wave-induced horizontal
        mean flow (cite)
        \begin{equation}
            \bar{U}_0 \equiv \ev*{u_x}_x = \frac{\ev*{u_xu_z}_x}{c_{gz}}
        \end{equation}
        We've denoted $c_{gz} = \at{\pd{\omega}{k_{z}}}_{k_{z0}}$ the vertical group
        velocity. In terms of our forcing amplitude,
        \begin{equation}
            \bar{U}_0(F) = -\frac{A(F)^2}{2}e^{\frac{z - z_0}{H}} \frac{k_z}{k_x}c_{gz},
                \label{eq:u0_lin}
        \end{equation}
        where $c_{gz} = -\frac{Nk_{0x}k_{0z}}{\p*{k_{0x}^2 + k_{0z}^2 +
        \frac{1}{4H^2}}^{3/2}}$.
\end{itemize}
% Linear solution: e^z/2H, dispersion relation, phase/group velocity. Observe is
% exact solution of nonlinear.
% Linear solution is globally unstable to RTI though, thus not stable
% equilibrium of IVP
%
% Nonlinear: Critical layer absorption (viscous = Hazel, nonlin = Winters),
% wave-induced mean flow, horizontal momentum flux

@@

As not to interfere with the incompressibility constraint, we force the system
on the density equation, replacing \autoref{eq:fc_density} with
\begin{equation}
    \md{\rho}{t} = Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
            \cos \p*{k_{0x}x - \omega_0 t}.
\end{equation}
Using a narrow Gaussian profile excites a broad $k_z$ wavenumber spectrum, and
only the $k_{0z}$ satisfying the IGW's dispersion relation for the given
$k_{0x}, \omega_0(k_{0x}, k_{0z})$ will propagate. The fluid equations then take
form
\begin{subequations}\label{se:vol_drive}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \md{\rho}{t} &= Fe^{-\frac{(z - z_0)^2}{2\sigma^2}}
                e^{i(k_xx - \omega t)},\\
        \md{\vec{u}}{t} + \frac{\vec{\nabla}P}{\rho}
            + g\hat{z} &= 0.
    \end{align}
\end{subequations}

The linearized \autoref{se:vol_drive}, obtained by dropping the advective term
in the material derivative, can be solved analytically up to good accuracy. We
may first approximate the driving term using $e^{\frac{-(z - z_0)^2}{2\sigma^2}}
\approx \sqrt{2\pi \sigma^2}\delta(z - z_0)$ if $\sigma \to 0$\footnote{$\sigma$
must be large enough to be numerically resolved by the spectral code.}. This
system is solved exactly by matching the two homogeneous solutions above and
below $z_0$, We may then approximately relax the solution to $k_{0z}\sigma
\lesssim 1$: an extra factor of $e^{-\frac{(k_{0z}\sigma)^2}{2}}$ arises
compared to the $\delta$-function solution (evaluating the Fourier Transform
of $e^{-\frac{(z - z_0)^2}{2\sigma^2}}$ at $k_z = k_{0z}$), and we obtain
\begin{align}
    A(F) ={}& \frac{Fgk_{0x}^2}{\rho_0(z_0)\omega_0^2}
        \frac{1}{2ik_z}\frac{e^{\frac{-(k_{0z}\sigma)^2}{2}}}
        {\sqrt{2\pi\sigma^2}},\\
    u_{1z}(x, z, t) ={}& A(F) \times\nonumber\\
        &{}\begin{cases}
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x + k_{0z}(z - z_0) - \omega_0 t)}
            & z > z_0\\
        e^{\frac{z - z_0}{2H}}e^{i(k_{0x}x - k_{0z}(z - z_0) - \omega_0 t)}
            & z < z_0\\
    \end{cases}.\label{eq:uz_lin}
\end{align}

@@

We nondimensionalize by taking $H = N = \rho_{ref} = 1$

%%%
These are implemented as:
\begin{subequations}\label{se:curr_num}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\Upsilon}{t} - \frac{u_z}{H}
            - \nu \nabla^2 \Upsilon &= -\Gamma(z) \Upsilon
                - \p*{\vec{u} \cdot \vec{\nabla}}\Upsilon
                + \frac{F}{\rho_0(z)}e^{-\frac{(z - z_0)^2}{2\sigma^2}}
                    \cos \p*{k_xx - \omega t},\\
        \pd{u_x}{t} + \pd{T}{x} + gH\pd{\Upsilon}{x}
            - \nu \nabla^2 u_x
            &= -\Gamma(z) u_x
                - \p*{\vec{u} \cdot \vec{\nabla}}u_x
                - T_1 \pd{\Upsilon}{x},\\
        \pd{u_z}{t} + \pd{T}{z} + gH\pd{\Upsilon}{z} - \frac{T_1}{H}
            - \nu \nabla^2 u_z &= -\Gamma(z) u_z
                - \p*{\vec{u} \cdot \vec{\nabla}}u_z
                - T_1 \pd{\Upsilon}{z},\\
    \Gamma(z) &= 7.5\s*{2 + \tanh \frac{z - z_T}{(L_z - z_T) / 2}
        + \tanh \frac{z_B - z}{z_B / 2}},\label{eq:Gamma}
    \end{align}
\end{subequations}
where $z_B = 0.05L_z, z_T = 0.95L_z$ are the boundaries of the damping zones. A
Navier-Stokes numerical viscosity $\nu$ is used to damp high wavenumbers and
regularize the nonlinear cascade at near grid resolution: $\nu \sim 0.2
\frac{\omega}{\abs*{k_z}}\frac{L_z}{2\pi N_z}$ was found to be suitable for $N_z
= 1024$, where $k_z$ is the wavenumber of the excited linear mode.
%%%

@@

The exact implementation of these equations in Dedalus is described in
\autoref{ss:strat_impl}.

@@

\section{Equation Implementations}\label{ss:strat_impl}

We denote $x \in [0, L_x], z \in [0, L_z]$ the simulation domain and $N_x, N_z$
the number of spectral modes in the respective dimensions. We perform direct
numerical simulation of \autoref{se:vol_drive} with the open-source
pseudo-spectral code Dedalus (CITE).

Numerically, the nonlinear $\frac{\vec{\nabla}P}{\rho}$ term is problematic: we
desire a system where the fluid fields are not divided by one another. We
introduce $\varpi = \frac{P}{\rho}$ instead, then mandate $\rho_0, \varpi_0$
background fields satisfy hydrostatic equilibrium $\vec{\nabla}\varpi_0 +
\varpi_0 \vec{\nabla}\rho_0 + g\hat{z} = 0$. Taking isothermal stratification,
we find $\varpi_0 = gH$. We further change variables to $\Upsilon = \ln \rho -
\ln \rho_0$ and $\varpi_1 = \varpi - \varpi_0$ deviations from the background
state to obtain a system of equations at most quadratic in fluid fields:
\begin{subequations}\label{se:fc_var}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \pd{\Upsilon}{t} + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon
            - \frac{u_z}{H} &= 0,\\
        \pd{u_x}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_x
            + \pd{\varpi_1}{x} + gH\pd{\Upsilon}{x}
            + \varpi_1 \pd{\Upsilon}{x} &= 0,\\
        \pd{u_z}{t} + \p*{\vec{u} \cdot \vec{\nabla}}u_z
            + \pd{\varpi_1}{z} + gH\pd{\Upsilon}{z}
            + \varpi_1 \pd{\Upsilon}{z} - \frac{\varpi_1}{H} &= 0.
    \end{align}
\end{subequations}
It bears noting that these equations are exactly equivalent to the original
Euler equations and hence conserve horizontal momentum.

\subsection{Artificial Dissipation}

The nonlinear terms in the above equations will transfer energy from lower
wavenumbers to higher wavenumbers. Since spectral codes have no numerical
dissipation, artificial dissipation must be added. To ensure the dissipitive
system conserves horizontal momentum, we begin by adding dissipitive terms to
the flux-conservative form of the Euler fluid equations \autoref{se:fc_orig} (we
use stress tensor $\tau_{ij} = P\delta_{ij}$):
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \rho + \vec{\nabla} \cdot (\rho \vec{u} - \nu
            \vec{\nabla}(\rho - \rho_0)) &= 0,\label{eq:visc_cons_mom}\\
        \partial_t (\rho \vec{u}) + \vec{\nabla} \cdot (\rho \vec{u} \vec{u} +
            \mathrm{diag}(\rho \varpi) - \nu \rho \vec{\nabla}\vec{u}) + \rho g
            \hat{z} &= 0.
    \end{align}
\end{subequations}
The same $\nu$ is used for both the diffusive and viscous term, though this is
not required. Since the dissipation is not physical and is purely used for
numerical stability, we choose it such that hydrostatic equilibrium is not
modified. Some algebraic manipulation to re-cast it in the form of
\autoref{se:fc_var} gives
\begin{subequations}
    \begin{align}
        \vec{\nabla} \cdot \vec{u} &= 0,\\
        \partial_t \Upsilon + \p*{\vec{u} \cdot \vec{\nabla}} \Upsilon -
            \frac{u_z}{H} - \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}} &= 0,\\
        \partial_t \vec{u} + \p*{\vec{u} \cdot \vec{\nabla}}\vec{u} +
            \vec{\nabla} \varpi + \varpi \vec{\nabla} \Upsilon - \nu \nabla^2
            \vec{u} + \vec{u} \nu\p*{\nabla^2 \Upsilon + \p*{\vec{\nabla}
            \Upsilon} \cdot \p*{\vec{\nabla}\Upsilon} - \frac{2}{H}\partial_z
            \Upsilon + \frac{1 - e^{-\Upsilon}}{H^2}}&{}\nonumber\\
        - 2\nu \p*{\p*{\p*{\vec{\nabla}\Upsilon} \cdot \vec{\nabla}}\vec{u} -
            \frac{1}{H}\partial_z \vec{u}} - \frac{\varpi_1}{H} &= 0.
    \end{align}
\end{subequations}
Hydrostatic equilibrium is still $\vec{\nabla} \varpi_0 + g\hat{z} = 0$ where
$\rho = \rho_0, \vec{u} = 0$. Including the damping layers and forcing terms as
described in \autoref{ss:numerics}, we finally obtain the full system of
equations as simulated in Dedalus:
